Measurement of a dwarf galaxy rotational velocity from its Lyman-alpha line emission

### 2016-09-22 paper/letter.tex
updated references and text

A dynamical mass of $10^{9} M_{\odot}$ in a sphere of 2 kpc could be
accomodated by a dark matter halo of at least $10^{13}$
M$_{\odot}$ in mass \cite{2011ApJ...726..108T}, which
seems unlikely as it would mean that  \tol\ is at the center of a
galaxy group.
The remaining possibility is that \tol\ hosts a supermassive black hole of
at least $10^{9} M_{\odot}$.

@@

In order to map out this large parameter space, we randomly drew
$2500$ sets of parameters within a observationally realistic range
(based on the considerations of \cite{Laursen2013ApJ...766..124L})
yielding a large variety of single-, double- and triple-peaked
spectra. 
The full analysis of the the spectral features as well as
more details on the radiative transfer are presented in
\cite{Gronke2016}.

For the current work, we computed the $\chi^2$ for each of the $2500$
Furthermore, we found that some parameters such as the magnitude of
the random clump motion $\sigma_{\rm cl}$ improved the fit
significantly whereas others did not.

We find that the best constrained parameters are the clump velocity
dispersion $\sigma_{\rm cl}=$\sigmaclump,  the outflowing clump
velocity $v_{\infty {\rm, cl}}=$\inftyclump\ and the probability that
the \lya\ emission comes from the clumps $P_{\rm cl}=$\probaclump.

@@

\section*{Physical Interpretation}

@@

From the multiphase model we obtained $v=$\sigmaclump
and $r=5$ kpc. This corresponds to a dynamical mass of $M_{\rm
  dyn}=2.8^{+1.3}_{-1.2}\times 10^{9}$ $M_{\odot}$ which is at least
$5$ times the HI mass plus the stellar mass estimated from
observations.

In the rotation model the size of the spherical region, $r_{s}$, can
be inferred from the relationship $\tau = \sigma_0 n r_s$, where
$\sigma_0=5.898\times 10^{-14}$cm$^{-2}$ is the Lyman$\alpha$
cross section at the  line's center, $n$ is the number density. 
This expression can be rewritten as
\begin{equation}
r_{s} = 0.055 \left(\frac{\tau}{10^7}\right) \left(
\frac{\mathrm{atoms\ cm^{-3}}}{n}\right) \mathrm{kpc}.
\end{equation}
%
Using this result and assuming a bound for the number density of
$0.01<n/\mathrm{atoms\ cm^{-3}}<0.1$ we have $0.55
< r_s/\mathrm{kpc}<5.5$.

This is consistent with the  constraint on the total HI mass and the
optical depth $\tau_0$ as we show next. The total HI mass can be
approximated as $M_H = \frac{4\pi}{3} m_H n r_s^3$, where
$m_H=1.67\times 10^{-24}$ gr is the mass of a single Hydrogen atom.
This allows us to write the total hydrogen mass as
\begin{equation}
M_{H} = 5.70 \times 10^{6} \left(\frac{\tau}{10^{7}}\right)
\left(\frac{r_s}{\mathrm{kpc}}\right)^2 M_{\odot}
\end{equation}
%
The upper limit from the HI mass non-detection imposes $r_{s} <
6.81$ kpc, which is consistent with the bound we found in the first
place, as we wanted to show.

Using those size bounds for the spherical model we find a constrain for
the dynamical mass of $5.45\times 10^{9}M_{\odot} < M_{\rm dyn} <
5.45\times 10^{10} M_{\odot}$.

>>>>

A dynamical mass of $10^{9} M_{\odot}$ in a sphere of 2 kpc could be
accomodated by a dark matter halo of at least $10^{13}$
M$_{\odot}$ in mass \cite{2011ApJ...726..108T}, which
seems unlikely as it would mean that  \tol\ is at the center of a
galaxy group.
The remaining possibility is that \tol\ hosts a supermassive black hole of
$10^{9} M_{\odot}$. This is two orders of magnitude higher
than the supermassive black hole found in the compact dwarf galaxy
M60-UCD1 \cite{2014Natur.513..398S}, which has a similar stellar mass
as \tol.

@@

In order to map out this large parameter space, we randomly drew
$2500$ sets of parameters within a observationally realistic range
(based on the considerations of \cite{Laursen2013ApJ...766..124L})
yielding a large variety of single-, double- and triple-peaked
spectra. 
The full analysis of the the spectral features as well as
more details on the radiative transfer are presented in
\cite{Gronke2016}.    
 For the current work, we computed the $\chi^2$ for each of the $2500$
models. We selected the best $50$ models with the lowest $\chi^2$. The
$\chi^2$ gap in those $50$ models is close to $3000$, the lowest
$\chi^2$ is close to $1200$. The total number of degrees of freedom is
$104$.

Then we performed a Kolmogorov-Smirnov test to compare each parameter
distribution in the best $50$ models against the parent distribution
of $2500$ models. 
If we obtain a p-value $<0.05$ for a given parameter, we conclude that
this parameter does influence the $\chi^2$ fit, as the distribution for
the best $\chi^2$ models is statistically different to the
distribution from the global sample of $2500$ models.

From this test we found that only three parameters influence the
$\chi^2$: the clump outflow velocity $v_{\infty,\rm{cl}}$ (p-value 
$10^{-18}$), the clump velocity dispersion $\sigma_{\rm cl}$ (p-value
$10^{-4}$) and the probability that the \lya\ emission comes from the
clumps $P_{\rm cl}$ (p-value $10^{-4}$).

The best values for those parameters that we report here correspond to
the values that produce the minimum
$\chi^2$. The 1-$\sigma$ uncertainty comes from a parabolic fit to the
$\chi^2$ as a function of $v_{\infty,\rm{cl}}$, $\sigma_{\rm cl}$,
$P_{\rm cl}$ around its corresponding minimum.
Under these conditions we find $\sigma_{\rm cl}=$\sigmaclump,
$v_{\infty {\rm, cl}}=$\inftyclump\ and $P_{\rm cl}=$\probaclump.

@@

\section*{Dynamical Mass Estimates}

@@

As described in the main the letter, we use the 3D half-luminosity
radius, $r_{s}$, as the typical size for the HI region.

%From the multiphase model we obtained $v=$\sigmaclump
%and $r=5$ kpc. This corresponds to a dynamical mass of $M_{\rm
%  dyn}=2.8^{+1.3}_{-1.2}\times 10^{9}$ $M_{\odot}$ which is at least
%$5$ times the HI mass plus the stellar mass estimated from
%observations.

%In the rotation model the size of the spherical region, $r_{s}$, can
%be inferred from the relationship $\tau = \sigma_0 n r_s$, where
%$\sigma_0=5.898\times 10^{-14}$cm$^{-2}$ is the Lyman$\alpha$
%cross section at the  line's center, $n$ is the number density. 
%This expression can be rewritten as
%\begin{equation}
%r_{s} = 0.055 \left(\frac{\tau}{10^7}\right) \left(
%\frac{\mathrm{atoms\ cm^{-3}}}{n}\right) \mathrm{kpc}.
%\end{equation}
%
%Using this result and assuming a bound for the number density of
%$0.01<n/\mathrm{atoms\ cm^{-3}}<0.1$ we have $0.55
%< r_s/\mathrm{kpc}<5.5$.

%This is consistent with the  constraint on the total HI mass and the
%optical depth $\tau_0$ as we show next. The total HI mass can be
%approximated as $M_H = \frac{4\pi}{3} m_H n r_s^3$, where
%$m_H=1.67\times 10^{-24}$ gr is the mass of a single Hydrogen atom.
%This allows us to write the total hydrogen mass as
%\begin{equation}
%M_{H} = 5.70 \times 10^{6} \left(\frac{\tau}{10^{7}}\right)
%\left(\frac{r_s}{\mathrm{kpc}}\right)^2 M_{\odot}
%\end{equation}
%
%The upper limit from the HI mass non-detection imposes $r_{s} <
%6.81$ kpc, which is consistent with the bound we found in the first
%place, as we wanted to show.

%Using those size bounds for the spherical model we find a constrain for
%the dynamical mass of $5.45\times 10^{9}M_{\odot} < M_{\rm dyn} <
%5.45\times 10^{10} M_{\odot}$.
